Hyperbolicity of Augmented Links in the Thickened Torus

### 2022-04-16 main.tex
made better twist region defn

We point out that \cite{kwon2020fully}, the first author
proved that \emph{fully} augmented links in the thickened torus
are hyperbolic, so this paper can be seen as a generalization
of that work.

@@

\begin{define}
Recall that a \emph{twist region}
in a link diagram is a maximal sequence of vertices
such that consecutive vertices are two vertices of a bigon face,
and consecutive bigons meet only at a vertex (not an edge).

TODO terrible defn.

We say that the \emph{direction} of a twist region
is the pair of faces (possibly the same face)
that are across from the bigons at the end;
if a twist region consists of only one vertex,
a direction is a pair of opposite faces (again, possibly the same face)
meeting at the crossing.
We say that a twist region consisting of only one vertex
is \emph{trivial}.
%%\footnote{There is an ambiguity of ``maximality'' in this definition:
%%for example, if a vertex of a link diagram meets exactly one bigon,
%%it can be considered as part of a twist region of length 2
%%(given by that bigon),
%%or it can be considered as part of a twist region of length 1,
%%as there are no bigons in the other ``direction''.
%%It should be clear which ``direction'' a twist region is going
%%when we deal with augmentations later.
%%}

For links in the thickened torus,
a \emph{twist region} in a link diagram of $L=\mathcal{L}/\Lambda$ in $\torus
\times I$, is the quotient of a twist region in the biperiodic link
$\mathcal{L}$.

A biperiodic link
$\mathcal{L}$ is called \emph{twist-reduced} if for any simple closed curve on
the plane that intersects the diagram of $\mathcal{L}$
transversely in four points, with two
points adjacent to one crossing and the other two points adjacent to another
crossing, the simple closed curve bounds a subdiagram consisting of a (possibly
empty) collection of bigons strung end to end between these crossings. We say
the diagram of $L$ is \emph{twist-reduced}
if it is the quotient of a twist-reduced biperiodic
link diagram.
\end{define}

We note that when a link diagram is cellular
(Definition \ref{d:cellular}),
a twist region in the torus cannot be a cycle;
otherwise, the face adjacent to the twist region would be an annulus.

@@

\begin{definition}
Let $D(K)$ be a twist-reduced diagram of a link $K$ in $\torus \times I$.
We define
\emph{augmenting} as a process in which an unknotted circle component,
called a \emph{crossing circle},
is added to one or more twist regions of $D(K)$
(see Figure \ref{fig:Augmentations});
we call the resulting link $L$ an \emph{augmented link obtained from $K$}.
We say $L$ is \emph{fully} augmented if $L$ is obtained by augmenting
$K$ at \emph{every} crossing/twist site.
\label{d:augmentation}
\end{definition}

@@

@@

If the faces of $G(\sT)$ are disks,
then $\sT$ can be decomposed into a union of pyramids,
obtained by coning each face to the cone point of $\sT$.
This also gives a decomposition of the corresponding ideal torihedron
$\sT^\circ$ into ideal pyramids.
We call these the \emph{pyramidal decompositions} of $\sT$ and $\sT^\circ$.

\begin{definition}
Let $G$ be a graph on the torus.
Let $v$ be a vertex and $e,e'$ be distinct edges that meet $v$.
A \emph{left (resp. right) bow-tie modification to $v,e,e'$}
is the process of removing $v,e,e'$ and adding in
a pair of triangular faces, which we refer to as a ``bow-tie''
(see \figref{f:bow-tie-modification}).
%as in Figure \ref{TODO} (a) (resp. (b)).
The edges of the bow-tie are of three types;
\emph{diagonal} edges, which do not touch $v$,
\emph{long} edges, which are ``parallel'' to the original edges $e,e'$,
and \emph{short} edges, which connect $v$ to the two new vertices.
We call the new triangular faces the \emph{bow-tie faces}.

\begin{figure}[ht]
\input{fig-bow-tie-modification.tikz}
\caption{Bow-tie modifications}
\label{f:bow-tie-modification}
\end{figure}

@@

\begin{definition}
\label{d:tori-decomp-graph}
Let $L$ be a link obtained from augmenting an alternating link $K$
with a cellular link diagram $D = D(K)$.
We define the \emph{top/bottom bow-tie graph of $L$} as follows.

@@

We now place a \emph{crossing edge} at each crossing of $K$,
connecting the top and bottom strands at the crossing
(see \figref{f:crossing-edges}).
We also three horizontal edges for each crossing circle
(see \figref{f:cut-slice-flatten}, leftmost diagram).
%(see Figure \ref{fig:crossingArc} (a)).
%so that for each crossing edge,
%one end of the edge lies on a bottom strand while the other
%end lies on a top strand as in Figure \ref{fig:crossingArc} (a).

\begin{figure} 
\input{fig-crossing-edges.tikz}
\label{f:crossing-edges}
\caption{Splitting and flattening crossing edges.}
%\centering 
%\begin{tabular}{cc}
%\includegraphics[width=4cm]{crossingArc.png}
%& \includegraphics[width=5cm]{crossingPush.png}
%\\
%(a)
%& (b)
%\end{tabular}
%\caption{(a) The black strands are
%part of the link and the red strand is the crossing edge.
%(b) The blue and
%red edges represent the split crossing edges and the shaded half disk is bounded
%by the crossing circle} 
%\label{fig:crossingArc} 
\end{figure}

%Next we construct the graph that will define the torihedra.
View the link from the point at infinity from the top end
($\torus \times \{1\}$) of the thickened torus.
At each crossing of $K$,
push the top strand towards the bottom strand,
splitting the crossing edge into two
identical edges and spreading them apart as in
\figref{f:crossing-edges}.
Note that after this operation, the link (which is no longer a link)
for both top and bottom look the same,
but a crossing edge associated to a crossing
is pushed in different directions;
this contributes to a ``$2\pi/n$'' twist
when gluing back the faces of the top and bottom torihedra.
%Figure \ref{fig:crossingArc}.
% (b).
%We push the link
%components to infinity and stretch the crossing edge so that we have flattened
%the link onto $\torus \times \{0\}$ except for the crossing circles,
%which remains perpendicular to the projection plane.

%Now place a disk on each crossing circle, so that the disk is bounded by
%the crossing circle.
Now for each crossing circle $c$, consider a spanning
(twice-punctured) disk $B_c$.
The following operations are depicted in \figref{f:cut-slice-flatten}.
$B_c$ intersects the projection plane $\torus \times \{0\}$,
cutting $B_c$ into two pieces $B_c^+,B_c^-$,
each being an ideal triangle.
We then slice along the disk $B_c$, turning it into two copies,
$B_c^{(1)},B_c^{(2)}$ (in no particular order);
each copy $B_c^{(j)}$ is also cut horizontally into two pieces,
$B_c^{(j),+},B_c^{(j),-}$.
We untwist all crossings in the twist region
which $c$ encircles,
rotating $B_c^{(j)}$ by $180^\circ$ for each crossing.
Then we flatten the disks;
the crossing circle is shrunk to a point,
as it is at infinity.

@@

More generally, if $K$ is as above with a twist-reduced diagram
containing bigons,
and $L$ is obtained by augmenting $K$
such that for every twist region with at least one bigon,
either the twist region is augmented,
or each vertex of the twist region is augmented
in the other direction,
or both.
%(and possibly other crossings),
then $L$ is hyperbolic.
\end{theorem}

@@

Now we have a decomposition of the two torihedra into
degenerate base-angled pyramids
(recall that that means some of the
interior dihedral angles $\theta^*$ are 0);
since we need the pyramids to be non-degenerate,
we modify the graph on the torihedra
and the angle assignments to make all $\theta^*$ nonzero
as follows.

@@

If $\bar{f}$ does not have vertices of the latter kind,
i.e. if $f$ does not meet short or long edges,
then we will not modify $f$.
So assume that $\bar{f}$ does have such a vertex,
and suppose it is right-augmented
(the other case is treated similarly).
Then for all right-augmented vertices $v_i$ of $\bar{f}$,
$f$ would meet the short,long edges $e_{i,0},e_{i,\pi}$,
while for all left-augmented vertices,
$f$ would meet the diagonal edges.
In particular,
the edges $e_{i,0}, e_{i,\pi}$ always appear in counter-clockwise order.

@@

We can get an angle-splitting for $(\Gamma_T',\theta^*)$,
using the angle-splitting for $\Gamma_T(L)$ for old edges,
and for a new edge $e$ (as in \figref{f:adding_edges})
that cuts through some face $f$ of $\Gamma_T(L)$,
we assign $\vphi_{\vec{e}} = \sum \vphi_{\vec{e'}}$,
where the sum is over all the edges $\vec{e'} \in \del f$
that on the other side of $\vec{e}$.
By non-degeneracy of the angle-splitting on $D'$,
these assignments to the new edges are all non-zero

>>>>

We point out that in \cite{kwon2020fully}, the first author
proved that \emph{fully} augmented links in the thickened torus
are hyperbolic, so this paper can be seen as a generalization
of that work.

@@

\begin{define}
Recall that a \emph{twist region}
in a link diagram in the plane is a maximal sequence of vertices
such that consecutive vertices are two vertices of a bigon face,
and consecutive bigons meet at exactly one vertex.
The \emph{length} of the twist region is the number of bigons.

We say a single vertex $v$ is a \emph{trivial twist region}
(or \emph{twist region of length 0})
if, for some cyclical ordering $f_1,f_2,f_3,f_4$
of the four faces adjacent to $v$,
$f_1,f_3$ are not bigons.
Note that if all four faces are not bigons,
then we think of $v$ as being a trivial twist region in two ways,
and in this sense, every vertex is part of exactly
two twist regions.
%
%We say that the \emph{direction} of a twist region
%is the pair of faces (possibly the same face)
%that are across from the bigons at the end;
%if a twist region consists of only one vertex,
%a direction is a pair of opposite faces (again, possibly the same face)
%meeting at the crossing.
%We say that a twist region consisting of only one vertex
%is \emph{trivial}.

For links in the thickened torus,
a \emph{twist region} in a link diagram of $L=\mathcal{L}/\Lambda$ in $\torus
\times I$, is the quotient of a twist region in the biperiodic link
$\mathcal{L}$. 
\end{define}

\begin{define}
A biperiodic link $\mathcal{L}$ is called \emph{twist-reduced}
if for any simple closed curve on
the plane that intersects the diagram of $\mathcal{L}$
transversely in four points,
with two points adjacent to one crossing
and the other two points adjacent to another crossing,
the simple closed curve bounds a subdiagram consisting of
a collection of bigons strung end to end between these crossings.
We say the diagram of $L$ is \emph{twist-reduced}
if it is the quotient of a twist-reduced biperiodic link diagram.
\end{define}

@@

\begin{definition}
Let $D(K)$ be a twist-reduced diagram of a link $K$ in $\torus \times I$.
We define
\emph{augmenting} as a process in which an unknotted circle component,
called a \emph{crossing circle},
is added to one or more twist regions of $D(K)$
(see Figure \ref{fig:Augmentations});
we call the resulting link $L$ an \emph{augmented link obtained from $K$}.
We say $L$ is \emph{fully augmented} if $L$ is obtained by augmenting
$K$ at \emph{every} crossing/twist site.
\label{d:augmentation}
\end{definition}

@@

We note that when a link diagram is cellular,
a twist region in the torus cannot be a cycle;
otherwise, the face adjacent to the twist region would
have non-trivial homology.

@@

Since the faces of $G(\sT)$ are disks,
$\sT$ can be decomposed into a union of pyramids,
where each pyramid is obtained by coning a face of $G(\sT)$
to the cone point of $\sT$.
This also gives a decomposition of the corresponding ideal torihedron
$\sT^\circ$ into ideal pyramids.
We call these the \emph{pyramidal decompositions} of $\sT$ and $\sT^\circ$.

\begin{definition}
Let $G$ be a graph on the torus.
Let $v$ be a vertex and $e,e'$ be distinct edges that meet $v$.
A \emph{bow-tie modification to $v,e,e'$}
is the process of removing $v,e,e'$ and adding in
a pair of triangular faces, which we refer to as
\emph{bow-tie faces} or a \emph{bow-tie}
(see Figure \ref{f:bow-tie-modification},
\ref{f:bow-tie-modification-general}).
The edges of the bow-tie are of three types;
\emph{diagonal} edges, which do not touch $v$,
\emph{long} edges, which are ``parallel'' to the original edges $e,e'$,
and \emph{short} edges.
%, which connect $v$ to the two new vertices.
%We call the new triangular faces the \emph{bow-tie faces}.
A bow-tie modification is a \emph{left} bow-tie modification
if, traveling from $v$ out along $e$ and $e'$,
the diagonal faces appear to the left of $e$ and $e'$,
respectively;
likewise for \emph{right} bow-tie modifications.

\begin{figure}[ht]
\input{fig-bow-tie-modification.tikz}
\caption{Bow-tie modifications and its relation to torihedra}
\label{f:bow-tie-modification}
\end{figure}

\begin{figure}[ht]
\input{fig-bow-tie-modification-general.tikz}
\caption{Another example of (left) bow-tie modification}
\label{f:bow-tie-modification-general}
\end{figure}

@@

\begin{definition}
\label{d:tori-decomp-graph}
Let $L$ be a link obtained from augmenting an alternating link $K$
with a cellular link diagram $D = D(K)$.
We define the \emph{top/bottom bow-tie graph of $L$} as follows.
(See \figref{f:tori-decomp-graph} for illustrations.)

@@

We now place a \emph{crossing edge} at each crossing of $K$,
connecting the top and bottom strands at the crossing
(see \figref{f:crossing-edges}).
We also three horizontal edges for each crossing circle
(see \figref{f:cut-slice-flatten}, leftmost diagram).

\begin{figure} 
\input{fig-crossing-edges.tikz}
\caption{Splitting and flattening crossing edges.}
\label{f:crossing-edges}
\end{figure}

View the link from the point at infinity from the top end
($\torus \times \{1\}$) of the thickened torus.
At each crossing of $K$,
push the top strand towards the bottom strand,
splitting the crossing edge into two
identical edges and spreading them apart as in
\figref{f:crossing-edges}.
Note that after this operation, the link (which is no longer a link)
for both top and bottom look the same,
but a crossing edge associated to a crossing
is pushed in different directions;
this contributes to a ``$2\pi/n$ twist''
when gluing back the faces of the top and bottom torihedra.

Now for each crossing circle $c$, consider a spanning
(twice-punctured) disk $B_c$.
The following operations are depicted in \figref{f:cut-slice-flatten}.
$B_c$ intersects the projection plane $\torus \times \{0\}$,
cutting $B_c$ into two pieces $B_c^+,B_c^-$,
each being an ideal triangle.
We then slice along the disk $B_c$, turning it into two copies,
$B_c^{(1)},B_c^{(2)}$ (in no particular order);
each copy $B_c^{(j)}$ is also cut horizontally into two pieces,
$B_c^{(j),+},B_c^{(j),-}$.
We untwist all crossings in the twist region
which $c$ encircles,
rotating $B_c^{(j)}$ by $180^\circ$ for each crossing.
Then we flatten the disks;
the crossing circle is shrunk to a point,
as it is at infinity.

@@

More generally, if $K$ is as above with a twist-reduced diagram
containing bigons,
and $L$ is obtained by augmenting $K$
such that for every twist region with at least one bigon,
either the twist region is augmented,
or each vertex of the twist region is augmented
in the other direction,
or both,
%(and possibly other crossings),
then $L$ is hyperbolic.
\end{theorem}

@@

Now we have a decomposition of the two torihedra into
degenerate base-angled pyramids
(recall that that means some of the
interior dihedral angles $\theta^*$ are 0);
since we need the pyramids to be non-degenerate,
we modify the graph on the torihedra
and the angle assignments to make all $\theta^*$ non-zero
as follows.

@@

If $\bar{f}$ does not have vertices of the latter kind,
i.e. if $f$ does not meet short or long edges,
then we will not modify $f$.
So assume that $\bar{f}$ does have such a vertex,
and suppose it is right-augmented
(the other case is treated similarly).
Then for all right-augmented vertices $v_i$ of $\bar{f}$,
$f$ would meet the short, long edges $e_{i,0},e_{i,\pi}$,
while for all left-augmented vertices,
$f$ would meet the diagonal edges.
In particular,
the edges $e_{i,0}, e_{i,\pi}$ always appear in counter-clockwise order.

@@

We can get an angle-splitting for $(\Gamma_T',\theta^*)$,
using the angle-splitting for $\Gamma_T(L)$ for old edges,
and for a new edge $e$ (as in \figref{f:adding_edges})
that cuts through some face $f$ of $\Gamma_T(L)$,
we assign $\vphi_{\vec{e}} = \sum \vphi_{\vec{e'}}$,
where the sum is over all the edges $\vec{e'} \in \del f$
that on the other side of $\vec{e}$.
By non-degeneracy of the angle-splitting on $D'$,
these assignments to the new edges are all non-zero.
